---
title: "Assingment 02"
author: "tony sulfaro"
date: "3/1/2019"
output: html_document
---

# load the libraries
```{r }
library(tidyverse)
```

# load the data
```{r }
music_data <- read_csv("out_data.csv")
```

Notice how all the dataframe types are col_character(). Lets fix that real quick

```{r}
#transform(music_data, timestamp = as.Date(timestamp))
```

# Add Additional Columns
```{r }
# ID column
# https://stackoverflow.com/questions/23518605/add-an-index-numeric-id-column-to-large-data-frame
#music_data <- tibble::rowid_to_column(music_data, "ID")
#music_data$day <- weekdays(as.Date(music_data$timestamp))

# date col to factor variable

# count times by day
```
# Questions:

How many listening sessions were 30 minutes or more?

What genre(s) of music do I listen to the most?

# Graph:
How many listening sessions were 30 minutes or more?
```{r}
ggplot(data = music_data) +
  geom_point(mapping = aes(x=timestamp, y=duration, colour = duration >= 30))+
  scale_color_manual(values = c("red", "blue"))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  geom_hline(yintercept = 30)+
  ggtitle("Listening Duration per Session") +
  xlab("Listening Session") +
  ylab("Total Minutes")
```

### Let's revise this graph a bit, try a stacked bar chart and group by per day instead of just per session.

```{r }
# aggregate geom bar
# https://stackoverflow.com/questions/40260048/geom-bar-tied-exactly-to-x-and-y-axis-without-aggregating
ggplot(data = music_data) +
  geom_bar(mapping = aes(x=day, y=duration, fill=music_genre),stat = "identity")+
  geom_hline(aes(yintercept=30), colour="#838B8B")+
  theme_minimal()+
  scale_x_continuous(breaks=seq(0,31,1))+
  scale_y_continuous(breaks=seq(0,200,10))+
  ggtitle("Total minutes listened by day") +
  labs(caption="Bars over the line indicate listening sessions over 30 minutes.")+
  xlab("Day of Month in January") +
  ylab("Total Minutes")
```

# Appendix

### Description
This dataset contains 12 days worth of data about my personal music listening habits. The data was recorded between the dates 2019-1-12 and 2019-1-24. There are several columns I recorded so lets take a look at them.

```{r}
str(music_data)
```

## Usage:
```{r}
music_data
```

## Format:

### Dimensions:
Number of rows = `r nrow(music_data)`

Number of columns = `r ncol(music_data)`

### Column definitions:
* year - year timestamp
* month - month timestamp
* day - day timestamp
* hour - hour timestamp
* minute - minute timestamp
* second - second timestamp
* music_genre - genre of music listened to
* task - what I was doing while listening to music
* location - where I was listening to music
* duration - how long I listened to music
* service - what service was used to listen to the music

# Essay

### Background:
The theme of this particular assignment was to apply what we learned during lab and use those methods an apply it to our dataset.

### Questions:
To help me better understand the listening habits I posed two questions:

* How long do I listen to music each day?
* What genre(s) of music do I listen to the most?

### Generating the Graph:
While reading in the data into a data frame is pretty cool from a technical standpoint, for the average user (and even technical ones) it is hard to derive meaning from looking at only the raw data. To solve this issue I created a visualization to bring out key trends in the data to help answer questions posed by myself.

To answer the question 'what days do I listen to music and how many days did I listen more than 30 minutes?' I decided that a stacked bar chart would be able to accomplish this when aggregating by day. This means if there were two listening sessions for a particular day then their durations would be totaled as one bar for that particular day. This also allowed me to add a horizontal line at duration = 30 to indicate which bars were higher or lower than that line to show which days I listened more or less than 30 minutes of music.


When answering 'what genre(s) of music do I listen the most?' Because I had already implemented a stacked bar graph by category of music listened to representing the categories was trivial because when setting the fill = category it shows clearly which categories were listened to for each particular day and for how long.
 
### Influence from Lecture 04:
Initially, I did not have much direction with how I was visualizing data. I used bar charts, pie charts, and treemaps to try and show which genre(s) were most popular or what days, in particular, I listened to music the most. While these visualizations from assignment 1 helped quantify the data, the colors used were confusing and didn't help to supplement the information.

Lecture 04 helped me use color meaningfully when constructing a graph. In this particular graph that I constructed each genre of music had its own color so that when it was included within a bar in the bar graph the viewer would be able to see how much time a particular genre listened to that day.

