---
title: "Assingment 03"
author: "tony sulfaro"
date: "4/5/2019"
output: html_document
---

# Setup
```{r, include=FALSE}
library(tidyverse)
library(dplyr)
library(fivethirtyeight)
```

# Activity 01

Let's take a look at the dataset documentation using ?comic_characters from the library fivethirtyeight.

We can see that the data originally came from a study conducted by 538 that examined comic books from DC and Marvel from ~1939 until 2014 (August 24th) by gender / publisher.

Some quick dataset dimensions: The dataset has `r nrow(comic_characters)` rows and `r ncol(comic_characters)` columns (variables).

A single row or entry in the dataset contains a character and various attributes associated with them like their first appearance, if they are alive or not, what their eye and hair color is, and their gender.

Here is a list of all the variables for each row in the dataset:

* publisher: Comic publisher
* page_id: Qnique ID for that characters page within the wikia
* name: Name of the character
* urlslug: url that takes you to the character page
* id: Identity status of the character (secret or public identity, Marvel only) (Eg: Spiderman is secret)
* align: If character is Good, Bad, or Neutral
* eye: Eye color of the character
* hair: Hair color
* sex: Sex of the character (male, female, etc)
* gsm: If the character is a gender or sexual minority (homosexual or bisexual)
* alive: Alive or deceased
* appearances: Number of times they appear in comic books. (as of Sep. 2, 2014)
* first_appearance: First month and year character appeared in a comic book
* month: month of characters first appearance, if availible
* year: year of characters first appearance, if availible
* date: date of characters first appearance, if availible

# Activity 02

Here is a list of some categorial variables in the dataset:

* publisher: Comic publisher
* name: Name of the character
* id: Identity status of the character (secret or public identity, Marvel only) (Eg: Spiderman is secret)
* align: If character is Good, Bad, or Neutral
* eye: Eye color of the character
* hair: Hair color
* sex: Sex of the character (male, female, etc)
* gsm: If the character is a gender or sexual minority (homosexual or bisexual)
* alive: Alive or deceased

After looking at these variables and the non-categorical ones some questions that I could ask could include:

* What is the proportion of each gender by comic book publisher?
* What are the most common attributes for a good character or bad one?
* Does the percentage of non-male characters increase over time?

# Activity 03

Let's generate some sweet visualizations to help us answer some questions! (Not those previous ones though, they were just sample questions to help me think about the data)

## Question 01

In	each	year	since	1961	and	for	both	publishers	combined,	what	percent	of	new	
characters	are	male	versus	female?

```{r}
comic_characters %>%
  filter(year >= 1961) %>%
  group_by(year, sex) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=year, y=count, color=sex)) + geom_line(stat = "identity",position="dodge")
```


## Question 02

What	proportion	of	good	versus	bad	characters	from	each	publisher	are	male	versus	
female,	since	1961?

```{r}
comic_characters %>%
  filter(year >= 1961) %>%
  count(year, sex) %>%
  ggplot(mapping = aes(x = year, y = sex)) +
  geom_tile(mapping = aes(fill = n))
```

## Question 03


# Activity 04

# Activity 05

