---
title: "Lab14 - Speed Dating Data and Joins"
author: "tony sulfaro"
date: "4/10/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
dates = read_csv("dates.csv")
people = read_csv("people.csv")
```

# Definitions 

## 01

Relational Data

Relational data is data that relates to one another from different sources. There can be two datasets that have seperate entries but they can be combined through joins to get more information.

## 02

Primary Key

A primary key is a specific entry in a dataset that can be used to resolve other entries in the dataset. Eg: given a timestamp it is possible to get a specific entry from the dataset by looking up the row with the timestamp and then looking across at other entries.

## 03

Mutating Join

Add new variables to a data frame by matching observations in with one another. For example, if someone were working with some flights data and they had in one table departures and arrivals from airports and one column had a tail number they could resolve tail number to a planes table and get more information about the plane (plane type, passenger capacity, etc.).

## 04

Filtering Join

This type of command filters entries in a dataframe based on weather or not they match an observation in a table. This is different because rather than resolve data and add it to a dataframe they are filtering records based on existence in another table. For example if there were incoming flights one could filter the rows to planes that are from a specific carrier.

# Describe the Data

## 05

Each observation in the dates file is a date that two people went on, iid is a person and pid is the partner id that they went on the date with.

## 06

There are relations between iid and pid to an entry from the people table. Each iid and pid can be used to lookup a person from the people table.

## 07

Primary key of dates table = iid (id of person on date)

Primary key of people table = iid (id of person on date that matches to pid in dates table)

## 08

```{r}
dates %>%
  count(iid) %>%
  filter(n >1)
```

## 09

Add surrogate key

```{r}
dates = dates %>% 
  mutate(surrogate.key = row_number())
dates
```

# Left Joins

## 10

```{r}
dates.people = dates %>%
 left_join(people, dates, by = c("iid" = "iid")) ## join by iid = iid (resolve person not participant to their attributes)
dates.people
```

## 11

```{r}
dates.people = dates.people %>%
  select(iid, round, position, order, pid, match, dec, attr, sinc,intel,fun,amb,shar,like,prob,met,match_es,race)
dates.people
```

## 12

```{r}
dates.people = dates.people %>%
 left_join(dates.people, dates, by = c("pid" = "iid")) ## resolve pid to date attributes, {attribute}.y is pid attributes
colnames(dates.people)
```

## 13

```{r}
dates.people = dates.people %>%
  select(iid,	pid,	round.x,	position.x,	match.x,	like.x, race.x, race.y)
dates.people
```

# Answer a Question using Relational Data

## 14

## 15

## 16

# Extra Credit

```{r}
people <- tribble(
 ~id, ~last.name, ~first.name,
 1, "Smith", "Alexander",
 2, "Williams", "Olivia",
 3, "Bell", "Emma",
 4, "Thompson", "Liam",
 5, "Garcia", "Sophia"
)
courses <- tribble(
 ~id, ~course.num, ~title, ~schedule, ~instructor, ~location,
~credits, ~prereqs,
 1, "MI220", "Understanding Users", "M W 3:00 PM - 4:20 PM", "Wash",
"233 Communication Arts Bldg", 3, NA,
 2, "MI250", "Intro Applied Programming", "M W 10:20 AM - 12:10 PM",
"Introne", "C134 Holden Hall", 3, NA,
 3, "MI320", "Reasoning with Data", "W 12:40 PM - 3:30 PM", "Rader",
"106 Farrall Ag Eng Hall", 3, "MI 220 and MI 250",
 4, "MI350", "Evaluating HCT", "Tu Th 10:20 AM - 12:10 PM", "Rader",
"233 Communication Arts Bldg", 3, "MI 220 and MI 250",
 5, "MI420", "Interactive Prototyping", "Tu Th 10:20 AM - 12:10 PM",
"Wyche", "C134 Holden Hall", 3, "MI 220 and MI 250"
)
people.courses <- tribble(
 ~person.id, ~course.id,
 1, 1,
 1, 2,
 2, 3,
 2, 4,
 3, 4,
 3, 5,
 4, 1,
 4, 2,
 5, 3,
 5, 4
)
```

After you	have	run	the	code	to	create	the	tables,	a	dplyr	command	that	uses	joins to	create another data	table	that	has	more	than	one	row	per	student,	and	the	columns:	last.name, first.name,	course.num,	title,	schedule,	instructor,	location,	credits,	prereqs. This	new	data table lists	which	courses	each	student	is	registered	for.

```{r}

```
