---
title: "Lab16 - NYC Data Deep Dive"
author: "tony sulfaro"
date: "4/26/2019"
output: html_document
---

# Question 01

```{r setup, include=FALSE}
library(tidyverse)
jobs_tibble = read_csv("week16lab_jobs.csv")
```

# Question 02

Yes the object is a tibble because when I type `jobs_tibble` into the console after reading it in it spits out a tibble of dimensions 5,584 x 6. It also lists column data types as well which is helpful, something that dataframes by default do not do.

# Question 03

No it is not a factor because after running this code chunk it returns False.

```{r}
is.factor(jobs_tibble$occupation)
```

# Question 04

```{r}
jobs_df = data.frame(jobs_tibble)
```

The new object is a dataframe because it returns all the rows instead of the first 10. You can also do `is.data.frame(jobs_df)` to have R check the type for you and it does return `r is.data.frame(jobs_df)`, indicating that it is a dataframe.


# Question 05

```{r}
another_tibble = as_tibble(jobs_df)
another_tibble
```

Let's convert the dataframe back to a tibble for fun. Using the command `as.tibbl(jobs_df)` this changes the dataframe back to a tibble. We can verify it with `is.tibble(another_tibble)` which returns `r is.tibble(another_tibble)`, meaning that it is a tibble once again.

# Question 06

Using the tibble that was converted to a dataframe and back to a tibble, print out 20 rows instead of the default 10.

```{r}
another_tibble %>%
  print(n=20, width = Inf)
```

# Question 07

Tibbles offer a little bit more control in how they are printed and stored within the R language. Tibbles give the user a little bit more flexibility to ovverride default behavior like printing.

# Question 08

```{r}
untidy_data = read_csv("week16lab_untidy_data.csv")
untidy_data
```

# Question 09

The data is not tidy because there is a variable column that should be split into seperate columns per variable. tsk tsk bad data storage right there folks. There are multiple entries for a single date which is not ok. There should be one date entry then an entry for each variable in the variable column.

# Question 10

```{r}
spread(untidy_data, key = variable, value = value)
```

# Question 11

One benefit of using datasets that are "tidy" is that they are easier to work with, rather than write complex functions to sift through variables in the "variable" column and match them up with values in the value column we can reference specific entries by looking at a single row instead of multiple.

# Question 12

```{r}
dogs_nyc = read_csv("week16lab_dogs_of_nyc.csv")
dogs_nyc
```

# Question 13

```{r}
dogs_nyc %>%
  group_by(AnimalName, BreedName, AnimalGender, AnimalBirthMonth, ZipCode) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

I created this dplyr command that essentially groups by various attributes that are associated with a dog, some nonunique ones like their name, BreedName, Gender, and the animal birth month. While it is possible that two dogs of the same breed of the same gender share the same birhday in the same zip code, it is unlikely that such an event would occur. By using these attributes we can use them as a key to get a unique pet entry in the dataset (like almost 100% of the time).

# Question 14

There are animal names that have "UNKNOWN" as the value as well as BreedName values that have "Unknown"

let's take a look at them using this

```{r}
missing_dogs = dogs_nyc %>% 
  filter(AnimalName == "UNKNOWN" | BreedName == "Unknown")
missing_dogs
```

it looks like there are `r nrow(missing_dogs)` missing values in either the AnimalName field or the BreedName field out of the `r nrow(dogs_nyc)` entries.


# Question 15

One concern when graphing the number of licenses when grouping by any attribute is the presence of duplicate entries in the dataset. Some people may have registered their pets more than once which would cause that pet to contribute to two or more places within the dataset instead of just one. One method would be to generate a key using a combination of attributes like Name, Breed, Gender, and BirthMonth to isolate individual animal entries and then graph them by Borough from there. This would eliminate duplicate entries and give a more representative depiction in the graph.

# EC

To make the data tidy Dr.Rader had to wrangle a speghetti mess of Excel sheets using the read_excel function within R. After getting the data loaded she had to manipulate the data and join it together all while asking questions on what the cell data was about and how it was supposed to be recorded. Dr. Rader was able to join seperate tabs of the Excel sheet into a single data source that was easier to build reporting off of. This included mutating the values within the sheet and reshaping it quite a bit from the looks of it.
